---
title: "ANOVA family"
author: "Amira Ibrahim"
date: "October 10, 2019"
output: html_document
---
#Attach Data
```{r HW8, echo=TRUE}
library(readr) 
HW8 <- read_csv("D:/Biostatistics/Lecture/Module 2/HW8.csv")
attach(HW8)
View(HW8)
```
##Check normality : 
```{r , echo=TRUE}
library(moments) 
skewness(FBS)
#accepted range from -1 to +1
kurtosis(FBS)
#accepted range from -2 to +2 may to +3
boxplot(FBS ~ Treatment , xlab = "Treatment" , ylab = "FBS" , main = "Effect of Drug on FBS")
#visually , the data is normally distributed
shapiro.test(FBS)
#P-value > 0.05 , So data in normally distributed 
```

#More than 2 levels in variable >> So use ANOVA family 

#1st case : phase III trial to assess the efficacy of single & double dosing of an oral hypoglycemic
# primary endpoint :acheving normal FBS 
# H0 : 
mu FBS(single dose) = mu FBS(double dose) = mu FBS(placebo)
```{r , echo=TRUE}
ANOVA1 <- (aov(formula = FBS ~ Treatment , data = HW8)) 
summary(ANOVA1)
#p-value (F distribution)  < 0.05  , reject H0
#there is significant difference >> The Drug has an effect on FBS
```

I need to Know which type of treatment is significant :
```{r , echo=TRUE}
TukeyHSD(ANOVA1)
plot(TukeyHSD(ANOVA1))
```
the best significant diffrence is between single dose & placebo

#2nd case : phase III factorial trial to assess the efficacy of single & double dosing of an oral hypoglycemic with or without vitamen B 
#primary endpoint :acheving normal FBS 
```{r , echo=TRUE}
ANOVA2 <- aov(formula = FBS ~ Treatment + `vitamin B` , data = HW8) 
summary(ANOVA2)
#there is significant diff by treatment on FBS  
#Vit.B  not  has any effect ( no significant diff)
```
#if Age is a prospected covariate in 1st case : 
```{r , echo=TRUE}
ANCOVA1 <- aov(FBS ~ Treatment + Age , data = HW8) 
summary(ANCOVA1)
#there is significant diff by treatment on FBS  
#Age  not  has any effect ( no significant diff)
```
#if Age is a prospected covariate in 2nd case : 
```{r , echo=TRUE}
ANCOVA2 <- aov(FBS ~ Treatment + `vitamin B` + Age , data = HW8) 
summary(ANCOVA2)
#there is significant diff by treatment on FBS  
#Age, vit.B  not  have  any effect ( no significant diff)
```

#3rd case : phase III trial to assess the efficacy of single & double dosing of an oral hypoglycemic
#(one way MANOVA)
#primary endpoint : acheving normal FBS & HbA1c 
```{r , echo=TRUE}
MANOVA1 <- manova(cbind(FBS,HbA1c)~Treatment , data = HW8)
summary(MANOVA1)
## there is significant diff , but I want to know the effect of treatment on each variable :
summary.aov(MANOVA1)

# there is significant diff (treatment , FBS), Drug has an effect on FBS  
# But Drug Not affect HbA1C 
```

#4th case : phase III factorial trial to assess the efficacy of single & double dosing of an oral hypoglycemic with or without vitamin B :
#(two way MANOVA)
#primary endpoint : acheving normal FBS & HbA1c 
```{r , echo=TRUE}
MANOVA2 <- manova(cbind(FBS,HbA1c)~Treatment + `vitamin B` , data = HW8)
summary(MANOVA2)
## to know the effect separately:
summary.aov(MANOVA2)
#Treatment has an effect on FBS only  
#Vit B not has any effect on FBS or HbA1c
```
#if Age is a prospected covariate in 3rd case :
```{r , echo=TRUE}
MANCOVA1 <- manova(cbind(FBS,HbA1c)~Treatment + Age , data = HW8)
summary(MANCOVA1)
## to know the effect separately:
summary.aov(MANCOVA1)
#Treatment has an effect on FBS only  
#Age  not has any effect on FBS or HbA1c 
```

#if Age is a prospected covariate in 4th case :
```{r , echo=TRUE}
MANCOVA2 <- manova(cbind(FBS,HbA1c)~Treatment + `vitamin B`+Age , data = HW8)
summary(MANCOVA2)
## to know the effect separately:
summary.aov(MANCOVA2)
#Treatment has an effect on FBS only  
#Both Vit.B & Age  not have any effect on FBS or HbA1c
```
